---
title: "Bayesian_Factorial_ANOVA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim of this markdown

The aim of this markdown is to implement a Bayesian factorial ANOVA in prepartion for analysis of my experiment 2 data. I will be using data from the *ToothGrowth* package. I will also be comparing this analysis to the Frequentist alternative.

## Load packages 

```{r load packages, message = FALSE}
# rm(list = ls()) - clear global environment
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(ggplot2) # data vis
library(BayesFactor) # performs Bayesian ANOVA
library(rstanarm) # bayesian regression modelling
library(brms) # bayesian modelling
library(bayestestR) # describes Bayesian models and posterior distributions
library(bayesplot) # allows plots for posterior predictive distributions
library(loo) # Bayesian model comparison
library(arm) # computes Bayes factor for glm model
library(insight) # get posterior parameters 
library(fitdistrplus) # investigating the distribution of my data
library(see) # helps with plotting for bayestestR package
library(emmeans) # marginal mean from Bayesian model
library(sjstats) # computes partial eta squared effect size
library(effsize) # computes cohen's D for follow up t-tests
library(Rmisc) # for confidence intervals on the interaction plots
library(lsr) # for calculating effect size 
```

## Load example data 

```{r cars}
data("ToothGrowth")
View(ToothGrowth)
```

The following data set is from an experiment with a 3 x 2 factorial design. The outcome varible is tooth length (*len*), one factor is supplement type (*vitamin C (VC) or orange juice (OJ)*) and the other is the dose of the supplement (*0.5, 1.0 and 2.0*).

Are there main effect of supplement and dose on tooth growth, and do the two interact? Firstly, lets analysis the data using frequentist methods.

## Summary statistics

```{r summary statistics}
ToothGrowth %>%
  dplyr::group_by(supp, dose) %>%
  dplyr::summarise(m = mean(len))
```

Summary statistics show that as dose increases, tooth length increases. This is the case for supplements.

# Interaction plots

```{r interaction plot with confidence intervals}
# computes standard error and 95% confidence intervals
CIerror_bar <- summarySE(ToothGrowth, measurevar = "len", groupvars = c("supp", "dose"))

ggplot(data = CIerror_bar, aes(x = dose, color = supp, group = supp, y = len)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = len - se, ymax = len + se), width = .1) +
  geom_point(data = ToothGrowth, aes(x = dose, y = len), alpha = 0.5) +
  ylab("Tooth length (mm)") +
  xlab("Dosage (mg)")
```

They appears to be a clear main effect of dose (as dose increases tooth length increases) and supplement (orange juice appears to provide increase tooth length). The interaction is less obvious. This I implement a 3x2 factorial ANOVA to investigate these effects. 

## Two way factorial ANOVA (frequentist)

```{r frequentist ANOVA}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

aov.result <- aov(len ~ supp * dose, data = ToothGrowth)
plot(aov.result, 2)

summary(aov.result)
etaSquared(aov.result)
```
The Q-Q plot plots theoretical quantiles agaist the standardised residuals to visualise normality. This plot appears to indicate that the residuals are normally distributed, thus the normality assumption of the ANOVA has not been violated. 

The 3x2 ANOVA reveals the following:

- A significant main effect of supplement with a medium effect size [f(1,56) = 12.32, p = 0.0008, eta^2 = 0.06]. Because there are only two levels, I can say that orange juice significantly improves tooth growth versus vitamin C. 

- A significant main effect of dose with a large effect size [f(1,56) = 133.42, p < 0.0001, eta^2 = 0.64]. This will need post hoc analyses to investigate. 

- A significant interaction between dose and supplement with a small effect size [f(1,56) = 5.33, p = 0.02, eta^2 = 0.03].

Thus with this interaction, I then perform post hoc comparisons to investigate this difference.

## Post hoc analyses of significant interaction

```{r tukeyHSD}
TukeyHSD(aov.result)
```

Tukey's HSD reveals that the main effect of dose comes about via the higher the dose, the bigger the increase in tooth growth.

The interaction of dose and supplement comes about via orange juice providing larger increases in tooth growth as the dose increases, however for the highest doses there was no difference in between orange juice and vitamin C. 

## Conclusions so far

Thus far, frequentist methods have revealed significnat main effects of dose and supplement. They have also revealed that the effect of dose significantly depends on the supplement - i.e. orange juice provides larger growth in small doses, but there is no difference in tooth growth between the supplements for the maximum doses. 

These conclusions are fair, and the correct analyses and corrections have been implemented. How do this compare with Bayesian equivalents?

## Fitting Bayesian model

```{r fitting Bayesian model}
post1 <- stan_glm(len ~ 1, data = ToothGrowth, 
                  family = gaussian(link = "identity"))
post2 <- update(post1, formula = . ~ supp)
post3 <- update(post1, formula = . ~ dose)
post4 <- update(post1, formula = . ~ supp + dose)
(post5 <- update(post1, formula = . ~ supp * dose))
```

This example is taken and adapted from here (http://mc-stan.org/rstanarm/articles/continuous.html#model-comparison). Here I fit a series of models. I set the family as gaussian with identity link function because the residuals are normally distributed:

- an intercept only model (when supplement and dose are 0, what is ToothGrowth length?). This acts as my "null" model.
- one for each predictor variable (supplement and dose)
- the two predictors combined
- the interacton between the two predictors

The benefit of this method is that I can use the "loo" package to perform a model comparison. This provides extra information regarding how and which predictors predict the data.

## Model comparison

```{r model compariosn for stan_glm models}
# Compare them with loo
loo1 <- loo(post1, cores = 2)
loo2 <- loo(post2, cores = 2)
loo3 <- loo(post3, cores = 2)
loo4 <- loo(post4, cores = 2)
loo5 <- loo(post5, cores = 2)
(comp <- loo_compare(loo1, loo2, loo3, loo4, loo5))
```

This has been taken from the example here (http://mc-stan.org/rstanarm/articles/continuous.html#model-comparison).

The *loo* and *loo_compare* functions allow me to compare the models. The *loo* method refers to the Leave-One-Out cross-validation. This method investigates the differences in predictive errors which is encapsulated by the *elpd_diff* (difference in predictive errors). 

Columns within the matrix are computed by generating pairwise comparisons between each model and the model with highest ELPD (i.e. highest predictive accuracy). Model 5 appears to be the best model - this contains the the two main effects and the interaction. However model 4 (both predictors) is not too far behind. To further investigate this, we can compute Bayes factor comparisons between the models.

## Bayes factor comparison between models

```{r Bayes factor comparison between models}

m1 <- brm(len ~ 1,
          data = ToothGrowth,
          family = gaussian(link = "identity"),
          save_all_pars = TRUE)

m2 <- brm(len ~ supp,
          data = ToothGrowth,
          family = gaussian(link = "identity"),
          save_all_pars = TRUE)

m3 <- brm(len ~ dose,
          data = ToothGrowth,
          family = gaussian(link = "identity"),
          save_all_pars = TRUE)

m4 <- brm(len ~ supp + dose,
          data = ToothGrowth,
          family = gaussian(link = "identity"),
          save_all_pars = TRUE)

m5 <- brm(len ~ supp * dose,
          data = ToothGrowth,
          family = gaussian(link = "identity"),
          save_all_pars = TRUE)

comparison_v_intercept <- bayesfactor_models(m2, m3, m4, m5, denominator = m1)
comparison_v_intercept

comparison_v_interaction <- bayesfactor_models(m1, m2, m3, m4, denominator = m5)
comparison_v_interaction
```

The Bayes factor comparison between each model comes from the following example (https://easystats.github.io/bayestestR/articles/bayes_factors.html).

First I fit the models using the *brms* package in order to extract the Bayes factors for model comparison. The first comparison is against the null (intercept only). This comparison shows that the data are all more likely (Bayes factors ranging from 29.63 to 9.18 x 10^17) under the comparison models than under the null model.

I then focus on the interaction model. The loo model comparison demonstrated that it was the best model however model 4 (both predictors without the interacion) was not that different. When I compare the Bayes factors, I find that the data are all much less likely under the null and individual main effect models than for the model containing main effects and interactions. However the Bayes factors comparison indicate that the data is no more likely under the model including the interaction than without it.

## Conclusions on Bayes factor model comparisons

This final comparison between model 4 and 5 is of particular interest. It contrasts the frequentist analysis, whereby a significant interaction effect (albeit with a small effect size) is identified and thus interpretted. Under a Bayesian analysis framework, the data is no more likely under the model including the interaction than without it, thus indicating a lack of evidence for it's influence on the data. 

## Investigating main effects of supplement and dose - 95% HDIs

```{r investigation into main effects}
posteriors_supp <- insight::get_parameters(post2)

ggplot(data = posteriors_supp) +
  geom_vline(xintercept = median(posteriors$suppVC), color = "red") +
  geom_density(aes(x = suppVC), fill = "red", alpha = 0.3) +
  geom_vline(xintercept = -.02, color = "black", size = 2) +
  geom_vline(xintercept = .02, color = "black", size = 2) +
  xlab("standardised effect size") +
  ggtitle("95% HDI for main effect of supplement")
  
rope_supp <- bayestestR::rope(post2, ci = 1, range = c(-0.02, 0.02))
rope_supp

posteriors_dose <- insight::get_parameters(post3)

ggplot(data = posteriors_dose) +
  geom_vline(xintercept = median(posteriors$dose), color = "blue") +
  geom_density(aes(x = dose), fill = "blue", alpha = 0.3) +
  geom_vline(xintercept = -.02, color = "black", size = 2) +
  geom_vline(xintercept = .02, color = "black", size = 2) +
  xlab("standardised effect size") +
  ggtitle("95% HDI for main effect of dose")

rope_dose <- bayestestR::rope(post3, ci = 1, range = c(-0.02, 0.02))
rope_dose
```

I compute the 95% HDIs for the main effects of supplement and dose. It is clear that for dose, 0 of the 95% most credible values fior the parameter value are practically equivalent to zero. For the supplement main effect, we see that 0.12% of the most credible values are practically equivalent to the null. These posterior distributions demonstrate that we should be very confident in the effect that is being observed for both main effects.

## Restriction models for the main effects

```{r restriction models}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

contrasts(ToothGrowth$dose) <- "contr.bayes"

post3 <- update(post1, formula = . ~ dose)

hyp_dose <- c("`0.5` < `1` & `1` < `2`")

# compute posterior marginal means
em_dose <- emmeans(post3, ~ dose)

# compute prior model estimated marginal means of this prior model
fit_dose_prior <- update(post3, prior_PD = TRUE)
em_dose_prior <- emmeans(fit_dose_prior, ~ dose)

bayesfactor_restricted(posterior = em_dose,
                       prior = em_dose_prior,
                       hypothesis = hyp_dose)
```


##TO DO

- Finish writing up main effects analyse for dose and supplement
- Plot stan_glm model using ggplot (http://mc-stan.org/rstanarm/articles/continuous.html#model-comparison)
- Read Bayesian stats paper
