---
title: "Estimating effect sizes for Bayesian models"
output: html_document
---

This R markdown will specify how to compute the effect sizes of Bayesian models. The first part will utilise the PlantGrowth data frame. I will then implement my own. I will utilise the rstanarm package. I will be following examples from the following online resource: (https://easystats.github.io/effectsize/articles/bayesian_models.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(ggplot2) # data vis
library(BayesFactor) # performs Bayesian ANOVA
library(rstanarm) # bayesian regression modelling
library(bayestestR) # describes Bayesian models and posterior distributions
library(bayesplot) # allows plots for posterior predictive distributions
library(loo) # Bayesian model comparison
library(arm) # computes Bayes factor for glm model
library(insight) # get posterior parameters 
library(fitdistrplus) # checking distribution
```

## Load data

```{r}
data("PlantGrowth")
```

Obtaining indices of effect for Bayesian models is currently an issue, but it can be done to an extent via one of two methods:

1) To compute standardised parameters.
2) To compute effect sizes via test statistic approximations.

## Implement contrasts and finding out distribution of data 

```{r}
descdist(PlantGrowth$weight, discrete = FALSE)

contrasts(PlantGrowth$group)

# setting my own contrasts - better addresses my research question
contrastmatrix <- cbind(c(-1, 1, 0), c(-1, 0, 1))

contrasts(PlantGrowth$group) <- contrastmatrix

contrasts(PlantGrowth$group)
```

Firstly I investigate the distribution of the data. This a quick and dirty way of doing this, but the descdist() function compares skewness and kurtosis of the data to different distributions. The cloest the data falls, the closer the fit of the data. Here is is very close to a normal distribution thus I can use the Gaussian family in my Bayesian model. 

I then set my contrasts. Because I have a control condition and 2 treatment conditions, I set my contrast matrix comparing the treatments to the control condition. 