---
title: "Bayesian analysis - draft write up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown document is my first attempt at writing up a Bayesian analysis of my data. The data I am using is from experiment 1. For this example, I will be investigating how heading angle affects RTs. For a reminder, both Threshold and Accumulator models propose reduced RTs as heading increases.

Firstly I analyse my data using a Bayesian ANOVA from the BayesFactor package. I then go on to use the rstanarm package to fit Bayesian generalised linear models to my data.

## Load packages 

```{r load packages, message =FALSE}
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(ggplot2) # data vis
library(BayesFactor) # performs Bayesian ANOVA
library(rstanarm) # bayesian regression modelling
library(bayestestR) # describes Bayesian models and posterior distributions
library(bayesplot) # allows plots for posterior predictive distributions
library(loo) # Bayesian model comparison
library(arm) # computes Bayes factor for glm model
library(insight) # get posterior parameters 
library(fitdistrplus) # investigating the distribution of my data
```

## Load data

```{r load my data}
# rm(list = ls()) - clear global environment
# home working directory
# setwd("C:/Users/Courtney/Documents/PhD/Project/Experiment_Code/experiment_1")
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/experiment_1")
temp = list.files(pattern = c("magnitudedata", "*.csv"))
myfiles = lapply(temp, read.csv)
magnitudedata <- do.call(rbind.data.frame, myfiles)
```

## Computing Bayesian ANOVA using BayesFactor package

```{r computing Bayesian ANOVA using BayesFactor}

magnitudedata <- magnitudedata %>%
  dplyr::filter(heading > 0)

magnitudedata$heading <- as.factor(magnitudedata$heading)

bf <- anovaBF(FirstSteeringTime ~ heading, data = magnitudedata)
bf

```

Using the anovaBF function tests the full hypothesis model versus the null model hypothesis. In my example, this will be the main effect of heading. 

The computed Bayes factor is 2.24 x 10^75. This indicates that the current data is 2.24 x 10^75 times more likely under the alternative hypothesis versus the null hypothesis i.e. there is a main effect difference between the levels of my heading factor.

However, am I testing the "right" hypothesis? Richard Morey (https://richarddmorey.org/2015/01/multiple-comparisons-with-bayesfactor-part-2-order-restrictions/#comment-7782) proposed that testing for differences between your factor levels refers to an *unconstrained model* (unconstrained in the sense that I am looking for any differences between my factor levels without specification). I actually have specific predictions on how my factor levels will differ. This refers to an *order constraint model*:

0.5 < 1.0 < 1.5 < 2.0

Already I have Bayes factor for the full model versus the null model. My order constraint model will allow me to make 2 more comparisons with my Bayes factor:

- Full model versus the order constraint model
- Order contraint model versus null model

Bayes factor is the degree of change from prior to posterior odds. Hence if we can compute prior odds, we can compute the Bayes factor. Because I have not specified a prior distribution, my prior odds have equal probability across my factor levels. With 4 factors, the prior odds are 1/8 as there are 8 orderings.

## Sampling and plotting posterior distribution

```{r sample from the posterior distribution}
samples <-  posterior(bf, iterations = 10000)
head(samples)

summary(samples[,2:5])

plot(samples[,2:3])
plot(samples[,4:5])
```

The *posterior* function allows me to sample from the posterior distribution of each heading.

I think the *mu* parameter refers to an estimated population mean parameter for reaction time. Each heading level is then compared to this mu parameter to compute the effect.

The plots show the posterior distributions for each heading level and the standardised effect size. For the *0.5 heading*, the HDI spans across positive effect sizes (i.e. reactions times are higher than *mu*). For *1.0 heading* the HDI spans across zero (reaction times are lower than *mu*). For *1.5 and 2.0 heading* the HDIs span negative standardised effects sizes (reaction times are lower than *mu*).

However these HDIs do how my order constraints...

## Creating order constraint model

```{r order constraint model}
## check order constraint
consistent = (samples[, "heading-1.5"] > samples[, "heading-2"]) &
  (samples[, "heading-1"] > samples[, "heading-1.5"]) &
  (samples[, "heading-0.5"] > samples[, "heading-1"])

N_consistent = sum(consistent)
```

I construct order constraints whereby I expect larger RTs for 1.5 than for 2, 1 than for 1.5, and 0.5 than for 1.

## Order constraint model versus full model

```{r constraint versus full model}
bf_contraint_against_full = (N_consistent / 10000) / (1 / 8)
bf_contraint_against_full
```

The posterior probability is the number of consistent order constraints from the posterior (*n_consistent*) divided by the number of samples (10000). Dividing the posterior by the prior distribution probabilites to get the Bayes factor of the constraint model versus the full model.

We generate a Bayes factor of 8. Hence the data has changed my opinion of the data, in favour of the constraint model, by a factor of 8. 

## Order constraint model versus null model

```{r constraint versus the null model}
## Convert bf1 to a number so that we can multiply it
bf_full_against_null = as.vector(bf)

## Use transitivity to compute desired Bayes factor
bf_restriction_against_null <- bf_contraint_against_full * bf_full_against_null
bf_restriction_against_null

```

If I multiple the Bayes factor from the full model by the Bayes factor of the contraint versus the full model, I can generate the Bayes factor for the constraint model versus the null model. This generates a Bayes factor of 1.79 x 10^76. This indicates the data is 1.79 x 10^76 more likely under the constraint model than under the null.

Overall, I can say that there is more evidence for constraint model versus the null than for the standard main effect model versus the null. However with a Bayes factor of 8, I would remain inconclusive on the constraint versus full model comparison.

## Conclusions so far

Using the anovaBF function gives me a nice introduction to Bayesian analyses. However even a Bayesian ANOVA relies on my error being normally distributed (which it is not). Because of this, I investigated Bayesian generalised linear models where I could account for my normal distribution of errors.

## Looking at the distribution of my data

```{r how is my data distributed}
fit_g  <- fitdist(magnitudedata$FirstSteeringTime, "gamma")

fit_ln <- fitdist(magnitudedata$FirstSteeringTime, "lnorm")

par(mfrow = c(2,2))
plot.legend <- c("gamma", "lognormal")
denscomp(list(fit_g, fit_ln), legendtext = plot.legend)
cdfcomp (list(fit_g, fit_ln), legendtext = plot.legend)
qqcomp  (list(fit_g, fit_ln), legendtext = plot.legend)
ppcomp  (list(fit_g, fit_ln), legendtext = plot.legend)
```

Here I looking at the distribution of my data.

*Log-normal distribution* is like a normal distribution but only contains positive values. This creates a skewed distribution with a right tail. These characteristics match my reaction time data, hence the good fit.

*Gamma distribution* is similar to *log-normal distribution* in that it is right skewed, and for continous data. 

For both of these distributions, the shape parameter is held constant whilst the scale parameter shifts. This is good for my data as I have a lot of variance in my responses. As both of these distributions are so simialr, I proceed with *Gamma*.

## Setting my contrasts

```{r construct contrast matrix}
magnitudedata <- magnitudedata %>%
  dplyr::filter(heading > 0)

magnitudedata$heading <- as.factor(magnitudedata$heading)

# default contrasts - orthogonal
contrasts(magnitudedata$heading)

# setting my own contrasts - better addresses my research question
contrastmatrix <- cbind(c(-1, 1, 0, 0), c(0, -1, 1, 0), c(0, 0, -1, 1))

contrasts(magnitudedata$heading) <- contrastmatrix

contrasts(magnitudedata$heading)
```

The default contrasts assume that the 0.5 heading condiiton is my control and compare each of my other headings against this. These are my "orthogonal"" contrasts in that they are linearly and statistically independent (each vector cannot be defined via a linear combination of the other vectors). 

However I did not feel that these were the best contrasts for my research question. I create a contrast matrix with 3 contrasts:

- 1.0 vs 0.5
- 1.5 vs 1.0
- 2.0 vs 1.5

These contrasts allow me to investigate whether RTs decrease as heading values increase. 

## Fitting Bayesian regression model

```{r fitting Bayesian GLM model, message = FALSE}
fit_glm <- stan_glm(FirstSteeringTime ~ heading, family = gaussian(), adapt_delta = 0.999, data = magnitudedata)
```

The stan_glm function requires a prior. I omitted the option of a prior and defaulted to the weakly informative ones. This is because 1) I do not have prior results on which to base my predictions and 2) the documentation indicates that for many applications, they will work well (https://www.rdocumentation.org/packages/rstanarm/versions/2.19.2/topics/priors)

I set adapt_delta arguement to *0.999* to reduce the stepsize during the model simulation. Stepsize specifies the resolution at which you are simulating through the posterior distribution. If the steps are too large, you might miss something and thus estimates become biased. This is known as divergence. The documentation proposes an adapt_delta arguement value such as this to reduce the chances of divergence (documentation divergence -> http://mc-stan.org/rstanarm/articles/aov.html)

For my first model, I specify a gaussian family for the distributions of my errors. This is not the case for my data, but I want to see the difference. 

## Summary of my Bayesian ANOVA model

```{r summarising the model}
summary(fit_glm)
mean(magnitudedata$FirstSteeringTime)
```
My Rhat values suggest the model succesfully converged.

*mean_ppd* in the estimates table can be used as a heuristic for how well the model describes the data. This value represents a best guess of the mean posterior predicitive distribution (distribution of simiulated data based on current data). Ideally this should be equal the mean of my outcome variable (RT). In this case, the values are quite similar.

## ROPE intervals 

```{r ROPE intervals and drawing from posterior distributions}
posteriors <- insight::get_parameters(fit_glm)

ggplot(data = posteriors) +
  geom_vline(xintercept = median(posteriors$heading1), color = "red") +
  geom_vline(xintercept = median(posteriors$heading2), color = "orange") +
  geom_vline(xintercept = median(posteriors$heading3), color = "blue") +
  geom_density(aes(x = heading1), fill = "red", alpha = 0.3) +
  geom_density(aes(x = heading2), fill = "orange", alpha = 0.3) +
  geom_density(aes(x = heading3), fill = "blue", alpha = 0.3) +
  geom_vline(xintercept = -.02, color = "black", size = 2) +
  geom_vline(xintercept = .02, color = "black", size = 2) +
  xlim(-.15, .15) +
  xlab("standardised effect size")

# establishing the percentage of the HDIs within the ROPE interval
percentage_in_rope_glm <- rope(fit_glm, ci = 1, range = c(-0.02, 0.02))

percentage_in_rope_glm
```

ROPE stands for "region of practical equivalence". Comparing your alternative hypothesis to a single point null does not make much sense theoretically, as it is very likely difference (even if it is extremely small). Thus we can define a ROPE which is an interval around your point null that is "practically equivalent" to the null point. 

Kruschke (2018) proposes that for linear models, ROPE intervals are defined as 0.1 (half of a small effect size) multiplied by the standard deviation of the outcome variable. Hence this is what is done here.

I can then draw from the posterior distributions and plot the HDIs for each of my contrasts. This provides a nice interpretation, as I can literally say:

"*95% of the the most credible value for standardised effect size are not practically equivalent to the null*". 

## Posterior predictive distributions

```{r posterior predictive distribution}
# observedresponses
y <- magnitudedata$FirstSteeringTime

# sample draws from the posterior distribution
yrep_aov <- posterior_predict(fit_glm, draws = 50)

color_scheme_set("brightblue")
ppc_dens_overlay(y, yrep_aov[1:50, ])
```

*Posterior distributions* explain unknown parameter (standardised effect size) and provides the 95% most credible values for that parameter. *Posterior predictive distributions* refer to a distribution of future predicted data based upon the data already seen. 

If the model is a good fit, you should be able to simulate similar data that was generated during the experiment.

My model demonstrates a reasonably good fit, however it under predicts the density of responses at 0.5 seconds. It also overestimates the peak of the responses (i.e. the highest density of responses) as slower than the observed peak responses.

## Computing a Bayes factor

```{r Bayes factor}

bf_glm <- bayesfactor_parameters(fit_aov, null = c(-.02,.02))
bf_glm

```
I can compute my Bayes factor for each contrast. This also allows me to compute my Bayes factor for the alternative hypothesis versus the ROPE interval null hypothesis, rather than a point null. This demonstrates substantial evidence for the alternative hypotheses.

This is based upon the example given here (https://easystats.github.io/bayestestR/articles/bayes_factors.html#testing-models-parameters-with-bayes-factors)

## Plotting regression line from Bayesian model

```{r plotting regression line}
draws <- as.data.frame(fit_glm)
colnames(draws)[1:2] <- c("a", "b")

ggplot(magnitudedata, aes(x = heading, y = FirstSteeringTime)) + 
  geom_point(size = 1, position = position_jitter(height = 0.05, width = 0.1), alpha = 0.3) + 
  geom_abline(data = draws, aes(intercept = a, slope = b), 
              color = "skyblue", size = 0.2, alpha = 0.25) + 
  geom_abline(intercept = coef(fit_glm)[1], slope = coef(fit_glm)[2], 
                   color = "skyblue4", size = 1) 
```

Here I plot the regression line from the Bayesian model. To plot uncertainty, I plot the estimated regression line at each draw from the postrior distribution. 

It is a poor fit to the data. This is likely due to me setting the family as *gaussian* when I know it should actually by *Gamma*.

From the example (http://mc-stan.org/rstanarm/articles/continuous.html#linear-regression-example)

## Bayesian model with gamma family

```{r Bayesian model with gamma}
fit_glm_gamma <- stan_glm(FirstSteeringTime ~ heading, family = Gamma(link = "identity"), adapt_delta = 0.999, data = magnitudedata)
```

## Re-plotting the regression line

```{r re-plotting model with gamma family}
draws_gamma <- as.data.frame(fit_glm_gamma)
colnames(draws_gamma)[1:2] <- c("a", "b")

ggplot(dplyr::filter(magnitudedata, heading <= 1), aes(x = heading, y = FirstSteeringTime)) + 
  geom_point(size = 1, position = position_jitter(height = 0.05, width = 0.1), alpha = 0.3) + 
  geom_abline(data = draws_gamma, aes(intercept = a, slope = b), 
              color = "skyblue", size = 0.2, alpha = 0.25) + 
  geom_abline(intercept = coef(fit_glm_gamma)[1], slope = coef(fit_glm_gamma)[2], 
                   color = "skyblue4", size = 1) 
```

Thus we see that when I properly model the distribution errors, the fit of the model to the data is much better.

## Model comparison using the loo package 

```{r model comparison using the loo package}
loo_aov <- loo(fit_aov)

loo_glm <- loo(fit_glm)

print(loo_compare(loo_aov, loo_glm), digits = 3)
```

**TO DO**
- start analysis by looking at the distribution of the data. BECAUSE it's gamma, I the choose stan_glm to fit the data

- then go through analysis and compute Bayes factor in the end - no point doing an ANOVA that assume normally distributed errors when I dont have them.

- compare this to a stan_glm with gaussian family just to show that the fit of my data is better with a gamma distribution

- remove regression line plots as they do not make sense with the contrasts i have
- could plot regression lines for the contrasts I do have though i.e. just for headings 0.5 and 1 for example... 

- send to Callum for commments 
