---
title: "Bayesian_restriction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The aim of this markdown file

This markdown details a description of how to implement a restriction model. The original model comes from the easy stats website (https://easystats.github.io/bayestestR/articles/bayes_factors.html#specifying-correct-priors-for-factors-with-more-than-2-levels) however I had issues with implementing my own example.

After reporting an issue with the bayestestR team, it became clear that there is a bug when trying to implement restriction models with a Gamma distribution. One of the package creates helped me formulate a solution around this bug, and this markdown details that solution also.

## Load packages 

```{r load packages, message =FALSE}
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(ggplot2) # data vis
library(BayesFactor) # performs Bayesian ANOVA
library(rstanarm) # bayesian regression modelling
library(bayestestR) # describes Bayesian models and posterior distributions
library(bayesplot) # allows plots for posterior predictive distributions
library(loo) # Bayesian model comparison
library(arm) # computes Bayes factor for glm model
library(insight) # get posterior parameters 
library(fitdistrplus) # investigating the distribution of my data
library(see) # helps with plotting for bayestestR package
library(emmeans) # marginal mean from Bayesian model
```

## Load my data and data used in the example

```{r load my data}
# rm(list = ls()) - clear global environment
# home working directory
# setwd("C:/Users/Courtney/Documents/PhD/Project/Experiment_Code/experiment_1")
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/experiment_1")
temp = list.files(pattern = c("magnitudedata", "*.csv"))
myfiles = lapply(temp, read.csv)
magnitudedata <- do.call(rbind.data.frame, myfiles)
```

## Setting heading levels as factors and setting contrasts

```{r factors}
magnitudedata <- magnitudedata %>%
  dplyr::filter(heading > 0)

magnitudedata$heading <- as.factor(magnitudedata$heading)
```

Here I remove the 0 heading condition and change my predictor variable to a factor. 

## Fitting Bayesian model for contr.sum contrast

```{r contr.sum model fit}
contrasts(magnitudedata$heading) <- "contr.sum"

fit_sum <- stan_glm(FirstSteeringTime ~ heading,
                      family = Gamma(link ="identity"),
                      adapt_delta = 0.999,
                      data = magnitudedata)
```

First I fit a Bayesian model under the default contrast setting (*contr.sum*). According to the example, this contrast setting will bias priors (which will thus bias Bayes factors for the restriction models).

## Order restriction and estimated marginal means

```{r order restriction}
hyp_sum <- c("`2` < `1.5` & `1.5` < `1` & `1` < `0.5`",
             "`0.5` > `2` & `0.5` > `1.5` & `0.5` > `1`")

# compute posterior marginal means
em_sum <- emmeans(fit_sum, ~ heading)

# compute prior model estimated marginal means of this prior model
fit_sum_prior <- update(fit_sum, prior_PD = TRUE)
em_sum_prior <- emmeans(fit_sum_prior, ~ heading)
```

First I set the restriction model. Each line is a restriction. To demonstrate the bias, I set two restrictions that I might be interested in:

- 2 < 1.5 < 1 < 0.5
- 0.5 > 2, 0.5 > 1.5, 0.5 > 1

I then compute my estimated marginal means from my fitted Bayesian model. These are the marginal means from the posterior distribution.

In order to get around the *Gamma* family bug, I need to set a prior-only model. Setting *prior_PD* as TRUE updates the original model but draws from the prior distribution. The *fit_sum_prior* object is equivalent to the *fit_sum* object but sampled from the priors. I then compute the estimated marginal means for this prior only model using the *emmeans* function. 

## Bayes factor for restricted model under contr.sum contrast

```{r resttriction for contr.sum contrast}
bayesfactor_restricted(posterior = em_sum,
                       prior = em_sum_prior,
                       hypothesis = hyp_sum)
```

Using the prior and posterior model, I can then calculate a Bayes factors for both of my restrictions as the Bayes factor details the degree of change from prior to posterior odds.

Automatically I can see an issue. Under the *contr.sum* contrast setting, prior probability odds are biased for some restrictions. In this example, one of the orders is *a priori* more likely than the other. This will thus influence the Bayes factor and the conclusions that are made between alternative and null hypotheses for these restrictions.

## Conclusions so far

Thus far I have detailed how the default *contr.sum* setting can bias prior for diffrent restrictions. The online example also explains how this is worsened under other contrast settings (*contr.treatment*). The only way around this is to compute orthogonal contrasts which are set via the *contr.bayes*.

According to the documentation (https://www.rdocumentation.org/packages/bayestestR/versions/0.4.0/topics/contr.bayes) the interpretation of the parameters becomes obscure when using this contrast. However it is essential for unbiased priors when calculating Bayes factors for restriction models. 

## Fitting Bayesian model for contr.bayes contrast

```{r contr.bayes model fit}
contrasts(magnitudedata$heading) <- "contr.bayes"

fit_bayes <- stan_glm(FirstSteeringTime ~ heading,
                      family = Gamma(link ="identity"),
                      adapt_delta = 0.999,
                      data = magnitudedata)
```

I change the contrast setting to *contr.bayes*. Doing this allows for better informed priors when dealing with 3 or more factor levels. It essentially ensures that each prior for every contrast is equal. Otherwise when it comes to the posterior distributions, a contrast with a slightly different prior will be biased and thus the Bayes factor will be altered. 

## Generate estimated marginal means for each heading from the fitted model

```{r estimated marginal means}
em_bayes <- emmeans(fit_bayes, ~ heading)
```

The *emmeans* function calculates the estimated margin means based from the posterior distribution of the Bayesian model fit.

## Creating a prior model is get around a bug in the bayestestR package

```{r prior model implementation}
fit_bayes_prior <- update(fit_bayes, prior_PD = TRUE)
em_bayes_prior <- emmeans(fit_bayes_prior, ~ heading)
```

First I update the model and set *prior_PD* as TRUE. This generates samples from the prior distributions for my parameters (i.e. the levels of my heading factor). This means that the *fit_bayes_prior* object is the same as the *fit_bayes* - it just draws from the prior distribution rather than the posterior distribution.

The *em_bayes_prior* object is then the estimated marginal means based upon the prior Bayesian model fit. 

## Setting the hypotheses for the restriction model

```{r hypothesis}
hyp <- c("`2` < `1.5` & `1.5` < `1` & `1` < `0.5`",
         "`0.5` > `2` & `0.5` > `1.5` & `0.5` > `1`")
```

If parameters are numbers, they need to be surrounded by back ticks when formulating the hypothesis. 

## Calculating Bayes factor for restriction model

```{r restriction Bayes}
bayesfactor_restricted(posterior = em_bayes,
                       prior = em_bayes_prior,
                       hypothesis = hyp)
```

Usually the *bayesfactor_restricted* function draws the prior probabilities from the *stan_glm* object. However, there is a bug with the Gamma distribution family in the bayestestR package. In order to get around this, I need to fit a prior only model using the prior predictive distribution. I then calculate estimated marginal means for this prior only model. I pass prior estimates as my priors arguement (as it is calculated from draws of the prior distribution) and the estimates from the orignal model as my posterior argument. With both prior and posteriors set, I can compute a Bayes factor (the degree of change from prior to posterior odds) based on my hypothesis.

This time, the difference between the priors is much smaller. The Bayes factors are much more reliable.

In my case however, is unlikely that I will be comparing restrictions so this might not be necessary. 
