---
title: "bayesain_gganimate"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file will detail how to visualise the accumulation of evidence that occurs during Bayesian analysis. Frequentist analyses assume that the null hypothesis model is true, and then compute the probability of the simulated data under the null being as extreme as the obtained data. If the probability is low, the null is rejected and the alternative hypothesis is accepted. 

## Load packages

```{r}
# rm(list = ls())
library(rstanarm)
library(ggplot2)
library(gganimate)
library(dplyr)
library(tidyr)
library(transformr)
library(gifski)
library(bayestestR)
library(see)
library(insight)
library(devtools)
install_github("easystats/estimate")
library(magick)
```

## Loading my own data

```{r}
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/experiment_1")
temp = list.files(pattern = c("magnitudedata", "*.csv"))
myfiles = lapply(temp, read.csv)
magnitudedata <- do.call(rbind.data.frame, myfiles)
```

```{r}
true_effect <- 0.5
prior_width <- 0.333
max_n <- 10


data <- bayestestR::simulate_correlation(n=max_n, r=true_effect)
vizmatrix <- estimate::visualisation_matrix(data["V2"])

# Initialize empty dataframes
priors <- data.frame()
posteriors <- data.frame()
params <- data.frame()
prediction <- data.frame()
correlation_data <- data.frame()
for(i in 5:nrow(data)){
  print(i)

  # Get data
  current_data <- data[1:i,]
  current_data$Evidence <- i
  correlation_data <- rbind(correlation_data, current_data)

  # Model and make predictions
  model <- rstanarm::stan_glm(V1 ~ V2,
                              prior = normal(0, prior_width),
                              data=current_data,
                              refresh = 0,
                              iter=10000,
                              chains=4,
                              warmup=4000)


  current_prediction <- estimate::estimate_link(model, length=50, data=vizmatrix, keep_draws=FALSE, draws=10)
  current_prediction$Evidence <- i
  prediction <- rbind(prediction, current_prediction)

  # Parameters
  posterior <- insight::get_parameters(model)$V2
  param <- bayestestR::describe_posterior(model, test=c("pd", "ROPE", "p_MAP"), rope_ci = 1)[2,]
  param$BF_null <- bayestestR::bayesfactor_parameters(model, verbose = FALSE)[2, "BF"]
  # param$BF_ROPE <- bayestestR::bayesfactor_rope(model, verbose = FALSE)[2, "BF"]

  param$Max <- bayestestR::density_at(posterior, median(posterior), method="KernSmooth")
  param$Evidence <- i
  params <- rbind(params, param)

  # Prior and posterior
  posterior <- bayestestR::estimate_density(posterior, method="KernSmooth")
  posterior$Evidence <- i
  posteriors <- rbind(posteriors, posterior)

  prior <- bayestestR::estimate_density(bayestestR::distribution_normal(1000, 0, prior_width), method="KernSmooth")
  prior$Evidence <- i
  priors <- rbind(priors, prior)
}


```

```{r}

p_correlation <- correlation_data %>%
  ggplot(aes(y=V1, x=V2)) +
  geom_point(size=3) +
  geom_ribbon(data=prediction, aes(ymin=CI_low, y=Median, ymax=CI_high, fill=Evidence), alpha=0.3) +
  geom_line(data=prediction, aes(y=Median, color=Evidence), size=1) +
  see::theme_modern() +
  xlab("Variable 1") +
  ylab("Variable 2") +
  scale_colour_gradient(low = "#FFC107", high = "#E91E63", guide = FALSE) +
  scale_fill_gradient(low = "#FFC107", high = "#E91E63", guide = FALSE) +
  gganimate::transition_time(Evidence) +
  labs(title = "Evidence (Sample Size): {frame_time}")




p_posterior <- posteriors %>%
  ggplot(aes(x=x, y=y)) +
  geom_area(data=priors, fill="#2196F3", alpha=1) +
  geom_segment(x = 0 , y = 0, xend = 0, yend = max(priors$y), size=0.5, color="#3F51B5", linetype = "dashed") +
  geom_segment(data=params, aes(x = Median , y = 0, xend = Median, yend = Max, color=Evidence), size=0.5, linetype = "dashed") +
  geom_line(aes(color=Evidence), size=1.5) +
  geom_vline(xintercept=true_effect, color="#E91E63", size=1) +
  scale_colour_gradient(low = "#FFC107", high = "#E91E63", guide = FALSE) +
  annotate("segment", x = 0.05, xend = true_effect, y = 1.25, yend = 1.25, colour = "#E91E63", size=1, arrow=arrow(length = unit(0.10, "inches"), type="closed")) +
  annotate("text", x = 0, y = 1.25, hjust = 1, colour = "#E91E63", label="True effect") +
  annotate("segment", x = -0.7, xend = -prior_width, y = 0.75, yend = 0.75, colour = "#2196F3", size=1, arrow=arrow(length = unit(0.10, "inches"), type="closed")) +
  annotate("text", x = -0.75, y = 0.75, hjust = 1, colour = "#2196F3", label="Prior") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 7)) +
  scale_y_continuous(expand = c(0, 0)) +
  see::theme_modern() +
  xlab("Effect") +
  ylab("Probability") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  coord_cartesian(xlim=c(-1, 1)) +
  gganimate::transition_time(Evidence)



p_significance <- params %>%
  select(Evidence, `p (direction)`=pd, `p (0)`=p_MAP, `p (ROPE)`=ROPE_Percentage, `BF (0)`=BF_null) %>%
  tidyr::pivot_longer(cols=-Evidence, names_to = "Index", values_to = "Value") %>%
  mutate(Index = forcats::fct_relevel(Index, "p (direction)", "p (0)", "p (ROPE)", "BF (0)")) %>%
  ggplot(aes(x=Evidence, y=Value, color=Index, group=1)) +
  geom_line(size=2) +
  scale_color_manual(values = c("p (direction)"="#9C27B0", "p (0)"="#f44336", "p (ROPE)"="#FFC107", "BF (0)"="#4CAF50", "BF (ROPE)"="#CDDC39"), guide=FALSE) +
  facet_wrap(~Index, nrow=5, scales = "free") +
  ylab("") +
  xlab("Evidence (Sample Size)") +
  see::theme_modern() +
  gganimate::transition_reveal(Evidence)


```

```{r}

p1 <- magick::image_read(animate(p_correlation, duration=nrow(data)/4, fps=20))
p2 <- magick::image_read(animate(p_posterior, duration=nrow(data)/4, fps=20))
p3 <- magick::image_read(animate(p_significance, duration=nrow(data)/4, fps=20))

```



Example data frame can be viewed 2 ways - each row is a trial or each row is a participant. As an example, I could use one participants and show the accumulation of evidence as the trials increases. Thus for the for loop would be 10:nrow(pNumDF) for 10 trials and so on.

I could do this with all trials in the magnitude dataframe however that might be computationally heavy. 



## Create example data

```{r}

prior_width <- 0.333

example <- magnitudedata %>%
  dplyr::select(heading, FirstSteeringTime, pNum) %>%
  dplyr::filter(pNum == 10)

vizmatrix <- estimate::visualisation_matrix(magnitudedata["heading"], length = 5)

# Initialize empty dataframes
priors <- data.frame()
posteriors <- data.frame()
params <- data.frame()
prediction <- data.frame()
correlation_data <- data.frame()

for(i in 10:nrow(example)){
  print(i)

  # Get data
  current_data <- example[1:i,]
  current_data$Evidence <- i
  correlation_data <- rbind(correlation_data, current_data)

  # Model and make predictions
  model <- rstanarm::stan_glm(FirstSteeringTime ~ heading,
                              prior = normal(0, prior_width),
                              data=current_data)


  current_prediction <- estimate::estimate_link(model, length=50, data=vizmatrix, keep_draws=FALSE, draws=10)
  current_prediction$Evidence <- i
  prediction <- rbind(prediction, current_prediction)

  # Parameters
  posterior <- insight::get_parameters(model)$heading
  param <- bayestestR::describe_posterior(model, test=c("pd", "ROPE", "p_MAP"), rope_ci = 1)[2,]
  param$BF_null <- bayestestR::bayesfactor_parameters(model, verbose = FALSE)[2, "BF"]
  # param$BF_ROPE <- bayestestR::bayesfactor_rope(model, verbose = FALSE)[2, "BF"]

  param$Max <- bayestestR::density_at(posterior, median(posterior), method="KernSmooth")
  param$Evidence <- i
  params <- rbind(params, param)

  # Prior and posterior
  posterior <- bayestestR::estimate_density(posterior, method="KernSmooth")
  posterior$Evidence <- i
  posteriors <- rbind(posteriors, posterior)

  prior <- bayestestR::estimate_density(bayestestR::distribution_normal(1000, 0, prior_width), method="KernSmooth")
  prior$Evidence <- i
  priors <- rbind(priors, prior)
}


```

## Make figures

```{r}

p_correlation <- correlation_data %>%
  ggplot(aes(y = FirstSteeringTime, x = heading)) +
  geom_point(size=3) +
  geom_ribbon(data=prediction, aes(ymin=CI_low, y=Median, ymax=CI_high, fill=Evidence), alpha=0.3) +
  geom_line(data = prediction, aes(y = Median, color = Evidence), size = 1) +
  see::theme_modern() +
  xlab("Variable 1") +
  ylab("Variable 2") +
  scale_colour_gradient(low = "#FFC107", high = "#E91E63", guide = FALSE) +
  scale_fill_gradient(low = "#FFC107", high = "#E91E63", guide = FALSE) +
  gganimate::transition_time(Evidence) +
  labs(title = "Evidence (Sample Size): {frame_time}")


p_posterior <- posteriors %>%
  ggplot(aes(x=x, y=y)) +
  geom_segment(x = 0 , y = 0, xend = 0, yend = max(priors$y), size=0.5, color="#3F51B5", linetype = "dashed") +
  geom_segment(data=params, aes(x = Median , y = 0, xend = Median, yend = Max, color=Evidence), size=0.5, linetype = "dashed") +
  geom_line(aes(color=Evidence), size=1.5) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 7)) +
  scale_y_continuous(expand = c(0, 0)) +
  see::theme_modern() +
  xlab("Effect") +
  ylab("Probability") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  coord_cartesian(xlim=c(-0.5, 0.5)) +
  gganimate::transition_time(Evidence)


p_significance <- params %>%
  select(Evidence, `p (direction)`=pd, `p (0)`=p_MAP, `p (ROPE)`=ROPE_Percentage, `BF (0)`=BF_null) %>%
  tidyr::pivot_longer(cols=-Evidence, names_to = "Index", values_to = "Value") %>%
  mutate(Index = forcats::fct_relevel(Index, "p (direction)", "p (0)", "p (ROPE)", "BF (0)")) %>%
  ggplot(aes(x=Evidence, y=Value, color=Index, group=1)) +
  geom_line(size=2) +
  scale_color_manual(values = c("p (direction)"="#9C27B0", "p (0)"="#f44336", "p (ROPE)"="#FFC107", "BF (0)"="#4CAF50"), guide=FALSE) +
  facet_wrap(~Index, nrow=5, scales = "free") +
  ylab("") +
  xlab("Evidence (Sample Size)") +
  see::theme_modern() +
  gganimate::transition_reveal(Evidence)

```

**Correlation dataframe**

In the correlation dataframe, there are 2 variabls and the investigation is how one affects the other. 

## Plotting figures

```{r}
p1 <- magick::image_read(animate(p_correlation, duration=nrow(example)/4, fps=20))
p2 <- magick::image_read(animate(p_posterior, duration=nrow(data)/4, fps=20))
p3 <- magick::image_read(animate(p_significance, duration=nrow(data)/4, fps=20))

final <- magick::image_append(c(p1[1], p2[1], p3[1]))
for(i in 2:length(p2)){
  combined <- magick::image_append(c(p1[i], p2[i], p3[i]))
  final <- c(final, combined)
}

# Save final
gganimate::anim_save("evidence_accumulation.gif", final)

```